### 2019-08-07 paper_I/table_cdips_targets.tex
update target table for page fit

 497093746702988672 &  2006361919000756992 &  2048442943413035776 &  5867618572762132864 \\
          Platais 3 &             ASCC 121 &           Teutsch 35 &           Loden\_1152 \\
           Dias2014 &             Dias2014 &             Dias2014 &       Kharchenko2013 \\
         799-012054 &           724-092321 &           630-074483 &            650278772 \\
            71.4458 &              337.708 &              294.081 &              209.832 \\
            69.6454 &              54.6607 &              35.8782 &             -59.2883 \\
            -5.1826 &             -3.48356 &             -3.19716 &              -3.6399 \\
          -0.478063 &             -1.28297 &             -2.04354 &             -3.93699 \\
           0.579745 &             0.196117 &             0.324896 &             0.397703 \\
            15.1735 &              15.4843 &              15.1863 &              16.4277 \\
            15.5786 &              16.3805 &              15.7505 &               16.991 \\
            14.6008 &              14.5227 &              14.4853 &              15.6807 \\
          Platais\_3 &             ASCC\_121 &             ASCC\_103 &           Loden\_1152 \\
          Platais\_3 &             ASCC\_121 &             ASCC\_103 &           Loden\_1152 \\
       string\_match &         string\_match &         string\_match &         string\_match \\
              False &                False &                False &                False \\
                NaN &                  NaN &                  NaN &                  NaN \\
                8.8 &                  6.4 &                 8.39 &                8.065 \\
                NaN &                  NaN &                  NaN &                0.092 \\