### 2019-08-07 paper_I/table_cdips_targets.tex
cdips target table now good

 Cluster &                                                 Platais 3 &             ASCC 121 &           Teutsch 35 &           Loden\_1152 \\
 Reference &                                                Dias2014 &             Dias2014 &             Dias2014 &       Kharchenko2013 \\
 Ext catalog name &                                       799-012054 &           724-092321 &           630-074483 &            650278772 \\
 RA &                                                        71.4458 &              337.708 &              294.081 &              209.832 \\
 Dec &                                                       69.6454 &              54.6607 &              35.8782 &             -59.2883 \\
 PMra &                                                      -5.1826 &             -3.48356 &             -3.19716 &              -3.6399 \\
 PMdec &                                                   -0.478063 &             -1.28297 &             -2.04354 &             -3.93699 \\
 Plx &                                                      0.579745 &             0.196117 &             0.324896 &             0.397703 \\
 $G$ &                                                       15.1735 &              15.4843 &              15.1863 &              16.4277 \\
 $G_{\rm Bp}$ &                                              15.5786 &              16.3805 &              15.7505 &               16.991 \\
 $G_{\rm Rp}$ &                                              14.6008 &              14.5227 &              14.4853 &              15.6807 \\
 K13 name match &                                          Platais\_3 &             ASCC\_121 &             ASCC\_103 &           Loden\_1152 \\
 Unique cluster name &                                     Platais\_3 &             ASCC\_121 &             ASCC\_103 &           Loden\_1152 \\
 How match &                                            string\_match &         string\_match &         string\_match &         string\_match \\
 Not in K13? &                                                 False &                False &                False &                False \\
 Comment &                                                       NaN &                  NaN &                  NaN &                  NaN \\
 K13 logt &                                                      8.8 &                  6.4 &                 8.39 &                8.065 \\
 K13 err logt                                                    NaN &                  NaN &                  NaN &                0.092 \\